---
title: "COVID-19 Canada Analysis"
author: "Liem Vu"
date: "2025-07-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
covid <- read_csv("covid_data.csv", show_col_types = FALSE)

canada <- covid %>%
  filter(country == "Canada", code == "CAN", continent == "North America") %>%
  select(date, total_cases, new_cases, total_deaths, new_deaths,
         people_vaccinated, people_fully_vaccinated, population)
glimpse(canada)
summary(canada)
```

1. Cases and Deaths
```{r}
canada_clean <- canada %>%
  mutate(date = ymd(date)) %>%
  drop_na(total_cases, total_deaths)

ggplot(canada_clean, aes(x = date)) +
  geom_line(aes(y = total_cases), color = "steelblue", linewidth = 1.2) +
  geom_line(aes(y = total_deaths), color = "firebrick", linewidth = 1.2) +
  labs(title = "Total COVID-19 Cases and Deaths in Canada",
       x = "Date", y = "Count") +
  theme_minimal()
```

Interpretation:
This line chart shows the growth of cumulative cases and deaths in Canada. Deaths increase more slowly than cases, likely due to vaccination and better treatment.

2. Vaccination Rate
```{r}
vaccine_data <- canada %>%
  mutate(date = ymd(date)) %>%
  drop_na(people_fully_vaccinated, population) %>%
  mutate(vaccine_rate = people_fully_vaccinated / population * 100)

ggplot(vaccine_data, aes(x = date, y = vaccine_rate)) +
  geom_line(color = "darkgreen", linewidth = 1.2) +
  labs(title = "Full Vaccination Rate Over Time (%)",
       x = "Date", y = "Percent") +
  theme_minimal()
```

Interpretation:
The percentage of fully vaccinated people rose sharply from early 2021, reaching over 70% in the end of the year.

3. Daily new cases and new deaths
```{r}
daily <- canada %>%
  mutate(date = ymd(date)) %>%
  drop_na(new_cases, new_deaths)

ggplot(daily, aes(x = date)) +
  geom_col(aes(y = new_cases), fill = "skyblue") +
  geom_col(aes(y = -new_deaths), fill = "tomato") +
  labs(title = "Daily New Cases (Blue) vs. Deaths (Red)",
       x = "Date", y = "Count") +
  theme_minimal()
```

Interpretation: 
The gap between new cases and new deaths widened after vaccines were rolled out. Deaths lagged cases by 2–3 weeks.